Computing with cut systems

### 2015-12-14 cutsystems-new.tex
Prior to changing Atiyahs to annihilation.

% Operators
\def\eval{\operatorname{ev}}
\def\res{\operatorname{Res}}
\def\Coker{\operatorname{Coker}}
\def\Ker{\operatorname{Ker}}
\def\im{\operatorname{Im}}
\def\can{\operatorname{can}}
\def\K{\mathbf{K}}
\def\D{\mathbf{D}}
\def\N{\mathbf{N}}
\def\LG{\mathcal{LG}}
\def\Ab{\operatorname{Ab}}
\def\Hom{\operatorname{Hom}}
\def\modd{\operatorname{mod}}
\def\Modd{\operatorname{Mod}}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\nN{\mathds{N}}
\def\nZ{\mathds{Z}}
\def\nQ{\mathds{Q}}
\def\nR{\mathds{R}}
\def\nC{\mathds{C}}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\tot}{Tot}
\DeclareMathOperator{\ch}{ch}
\DeclareMathOperator{\str}{str}
\DeclareMathOperator{\hmf}{hmf}
\DeclareMathOperator{\HMF}{HMF}
\DeclareMathOperator{\hf}{HF}
\DeclareMathOperator{\At}{At}
\DeclareMathOperator{\Cat}{Cat}
\DeclareMathOperator{\Spec}{Spec}

@@

\begin{abstract}
The bicategory $\LG$ of Landau-Ginzburg models has polynomials as objects and matrix factorisations as $1$-morphisms. The composition of $1$-morphisms naturally produces infinite rank matrix factorisations, which is a nuisance. In this paper we define a bicategory $\cat{L}$ which is equivalent to $\LG$ but for which composition of $1$-morphisms produces finite rank matrix factorisations equipped with the action of a Clifford algebra.
\end{abstract}

@@

\section{Introduction}

The bicategory of Landau-Ginzburg models $\LG$ has polynomials $W(x)$ called \textsl{potentials} as its objects (typically, the defining equation of an isolated singularity) and the $1$-morphisms $W(x) \lto V(y)$ are matrix factorisations of the difference $V(y) - W(x)$ over $\nC[x,y]$. This bicategory plays an important role in topological field theory \cite{McNameethesis,lgdual,genorb} and the study of homological knot invariants \cite{kr0401268}. In \cite{lgdual} it was shown that $\LG$ is a pivotal bicategory with adjoints, and closed formulas were given for the units and counits of adjunction. In this paper and its sequel we study an enrichment of $\LG$ which we refer to as a \emph{cut system}.

A finite rank matrix factorisation of $V(y) - W(x)$ is a matrix of polynomials
\[
D \in M_{2r}( \nC[x,y] ) \qquad \text{ satisfying } \qquad D^2 = (V(y) - W(x)) \cdot I_{2r}
\]
where we require that, dividing $D$ into $r \times r$ blocks, the upper left and bottom right blocks are zero matrices. The set of finite rank matrix factorisations is denoted $\LG(W,V)_{fin}$. For technical reasons the category of $1$-morphisms $\LG(W,V)$ contains, in addition to the finite rank matrix factorisations, those infinite rank matrix factorisations which are homotopy equivalent to finite rank ones; this means in particular that the inclusion
\[
\LG(W,V)_{fin} \lto \LG(W,V)
\]
is an equivalence of categories.

If we fix the size $r$ of the blocks and bound the degrees of the monomials appearing in $D$ by some integer $k$, we can parametrise the possible matrix factorisations by the closed points of an affine scheme $\mathscr{M}_{r,k}(W,V)$ whose coordinates are the coefficients of the various monomials in each row and column of $D$. Hence as sets we have
\begin{equation}\label{eq:param_frmf}
\LG(W,V)_{fin} = \bigsqcup_{r \ge 1} \bigcup_{k \ge 0} \mathscr{M}_{r,k}(W,V)_{\nC}\,.
\end{equation}
For various applications, two of which we sketch below, we would like to know that the basic operations of the bicategory $\LG$ respect this additional scheme structure; that is, that these operations are continuous in the coordinates which parametrise the $1$-morphisms. The most important and nontrivial such structure is the composition functor
\begin{gather*}
\LG(V,U)_{fin} \times \LG(W, V)_{fin} \lto \LG(W,U)\,,\\
(Y,X) \mapsto Y \circ X\,.
\end{gather*}
We would like therefore to produce a morphism of schemes (omitting indices)
\[
c: \mathscr{M}(V,U) \times \mathscr{M}(W,V) \lto \mathscr{M}(W,U)
\]
with the property that $c([Y], [X]) = [Y \circ X]$, where $[-]$ denotes the point of the parameter space corresponding to a given matrix factorisation.

The problem is that this doesn't make sense, because the matrix of polynomials defining the composite $Y \circ X$ contains monomials $x^i y^j z^k$ with $y$'s in them and we have no way to store the coefficients of such monomials in the space $\mathscr{M}(W,U)$. However one can prove that $Y \circ X$ is a summand in the homotopy category of a finite rank matrix factorisation (involving only $x$'s and $z$'s) and even describe this object and the corresponding idempotent explicitly \cite{dm1102.2957}. This is a finite amount of data which can be used to encode $Y \circ X$, so we just need to replace $\mathscr{M}(W,U)$ by some scheme which knows how to store such data.

\vspace{0.5cm}

After some refinements\footnote{See Remark \ref{remark:relation_to_toby_paper} for a discussion of why we end up encoding the finite model by Clifford actions rather than the idempotents of \cite{dm1102.2957}.} this amounts to replacing $\LG$ by an equivalent bicategory $\cat{L}$ with the same objects and (almost) the same $1$-morphisms, but whose composition operation is easily encoded as a morphism of schemes. We say \emph{almost} the same $1$-morphisms because in $\cat{L}$ the $1$-morphisms are still finite rank matrix factorisations, but now they may additionally equipped with the action of a Clifford algebra. It remains straightforward to parametrise these $1$-morphisms by the closed points of an affine scheme, but in contrast to $\LG$ the composition functor in $\cat{L}$ (which we refer to as the \emph{cut operation})
\begin{gather}
\cat{L}(V,U) \times \cat{L}(W,V) \lto \cat{L}(W,U)\,,\\
(Y,X) \mapsto Y \l X
\end{gather}
does indeed lift to a morphism of the parametrising schemes - fulfilling the original goal. The schemes parametrising $1$-morphisms in $\cat{L}$ together with the morphisms of schemes lifting composition, and vector bundles encoding the $2$-morphisms, is what we refer to as a \emph{cut system} in the sequel \cite{cutsystems2}. However the present paper is concerned only with defining the superbicategory $\cat{L}$ and proving that it is equivalent to $\LG$.

The paper is structured as follows: in Section \ref{section:background} we recall the definition of superbicategories and Clifford algebras, and introduce the Clifford thickening of a supercategory. In Section \ref{section:lg_cut_system} we define $\cat{L}$ and in Section \ref{section:equivllg} we prove it is equivalent to $\LG$ as a superbicategory without units. We do not consider unit $1$-morphisms here because $\cat{L}$ does not make working with units any simpler.

\subsection{Applications}

\subsubsection{Topological field theory}

The bicategory $\LG$ arises in topological field theory (TFT) as the data necessary to define a two-dimensional topological field theory with defects \cite{k1004.2307,dkr1107.0495}. It is expected that such a TFT should be associated to any suitably nice bicategory $\cat{B}$. In the formalisation of \cite[\S 3.1]{dkr1107.0495} the associated TFT is a symmetric monoidal functor
\[
\mathcal{Z}: \operatorname{Bord}_{2,1} \lto \operatorname{Vect}_\nC
\]
where the symmetric monoidal category $\operatorname{Bord}_{2,1}$ is defined roughly as follows: the objects are disjoint unions of (collared) circles labelled with objects and $1$-morphisms of $\cat{B}$, and the morphisms are diffeomorphism equivalence classes of two-dimensional manifolds with boundary and a labelled network of ``defects''.

\begin{figure}[bt]
$$
  \raisebox{-75pt}{\begin{picture}(300,130)
  \put(0,0){\scalebox{0.6}{\includegraphics{pic02.eps}}}
  \put(0,0){
     \setlength{\unitlength}{.55pt}\put(-22,-8){
%     \put(136,224)   {$ \phi_\text{in} $}
%     \put(140, 22)   {$ \phi_\text{in} $}
%     \put(433,142)   {$ \phi_\text{out} $}
	 \put(295,190)   {$ {\scriptstyle Y} $}
	 \put(240,190)   {$ {\scriptstyle X} $}
	 \put(335,85)   {$ {\scriptstyle Z} $}
	 \put(390,105)   {$ {\scriptstyle Q} $}
     \put(510,180)   {$ O_{3} $}
     \put( 10,230)   {$ O_{1} $}
     \put( 10,110)   {$ O_{2} $}
     }\setlength{\unitlength}{1pt}}
  \end{picture}}
$$
\caption{A world-sheet $M$ with defect lines from $\,(O_1,O_2)\,$ to
$\,(O_{3})$.}
\label{fig:wsh-defect}
\end{figure}

With $\cat{B} = \LG$ an example of a morphism is given in Figure \ref{fig:wsh-defect}. The two-dimensional regions of $M$ are labelled with potentials and the defect lines are labelled with $1$-morphisms in $\LG$, that is, matrix factorisations. In the special case where we use a single label $W(x)$ for all regions, the labels $X,Y,Z,Q$ for the defect lines are all chosen from the set of $1$-morphisms $W \lto W$. The value of $\mathcal{Z}$ on $M$ is a $\nC$-linear map
\be\label{eq:value_of_bordism}
\mathcal{Z}(M): \mathcal{Z}(O_1) \otimes \mathcal{Z}(O_2) \lto \mathcal{Z}(O_3)\,.
\ee
We would like to discuss an additional property of the physicist's Landau-Ginzburg model which is not encoded into the data of the functor $\mathcal{Z}$. Namely, it is a consequence of the way the defect conditions $X,Y,Z,Q$ enter the partition function (see for example \cite[(3.16)]{laz}) that the vector spaces $\mathcal{Z}(O_i)$ and linear map $\mathcal{Z}(M)$ should depend continuously on the coefficients defining the defect conditions (TODO of course not).

To represent this mathematically we might try to present $\mathcal{Z}(O_i)$ as a vector bundle over a scheme parametrising possible defect conditions and $\mathcal{Z}(M)$ as a morphism of these vector bundles. But there are some obstacles.
% \cite[(2.9)]{dkr1107.0495}

Consider the incoming circle $O_2$ with two points marked $X,Y$. The construction gives
\be\label{eq:zo2orig}
\mathcal{Z}(O_2) := \Hom_{\LG(W,W)}( \Delta_W, Y^{\vee} \otimes X) \cong H^0 \Hom_{\nC[x,x']}( Y, X )
\ee
which is the vector space of $2$-morphisms $Y \lto X$ in $\LG(W,W)$. Now suppose that the defect conditions are made to depend on parameters 
\[
Y = Y(t_1,\ldots,t_r)\,, \qquad X = X(u_1,\ldots,u_s)
\]
which satisfy their own equations and cut out a parametrising scheme $\mathscr{M}_Y \times \mathscr{M}_X \subseteq \mathbb{A}^{r+s}$. The question is: what kind of global object $\mathcal{Z}(O_2)$ on $\mathscr{M}_Y \times \mathscr{M}_X$ is formed by the finite-dimensional vector spaces
\[
\mathcal{Z}(O_2)_{P,Q} := H^0 \Hom_{\nC[x,x']}( Y(P), X(Q) )
\]
as $P$ varies over $\mathscr{M}_Y$ and $Q$ over $\mathscr{M}_X$? The problem is that \emph{a priori} we have
\be\label{eq:cohom_sucks}
H^* \Hom_{\nC[x,x',t,u]}( Y(t), X(u) ) \Big\vert_{t=P, u=Q} \neq H^* \Hom_{\nC[x,x']}( Y(P), X(Q) )\,.
\ee
%This is a particular case of a general phenomenon: any complex of $\nC$-vector spaces $C$ is isomorphic in the derived category $\mathbb{D}(\nC)$ to $\bigoplus_n H^n(C)[n]$, but over a general ring, for instance the coordinate ring of $\mathscr{M}_Y \times \mathscr{M}_X$, this is not the case. 
This discrepancy suggests that we should take as our global object $\mathcal{Z}(O_2)$ the Hom complex itself, as an object in the derived category\footnote{Note that it is not enough to take the derived category of coherent sheaves, because the Hom complex is made up of non-finitely-generated modules over $\mathscr{M}_Y \times \mathscr{M}_X$.}
\be\label{eq:wuhe20}
\mathcal{Z}(O_2) := \Hom_{\nC[x,x',t,u]}( Y(t), X(u) ) \in \mathbb{D}( \operatorname{qcoh} \mathscr{M}_Y \times \mathscr{M}_X )\,.
\ee
An alternative (and we believe simpler) approach is to take $\cat{L}$ rather than $\LG$ as the input to the definition of the TFT. Before doing so, let us observe that in the situation where $X,Y$ do not depend on parameters, we may equivalently\footnote{This definition is not quite equivalent, because the displayed composite produces a complex of vector spaces quasi-isomorphic to $H^* \Hom(Y,X)$ rather than $H^0 \Hom(Y,X)$, but this is not important here.} define the vector space $\mathcal{Z}(O_2)$ from \eqref{eq:zo2orig} via the following composite in $\LG$:
\be\label{eq:composite_intro_o2}
\xymatrix@C+2pc{
0 \ar[r]^-X & W(x) - W(x') \ar[r]^-{Y^{\vee}} & 0
}\,.
\ee
Now let $\mathcal{Z}'$ be the TFT defined using the bicategory $\cat{L}$ as input. Then $\mathcal{Z}'(O_2)$ may likewise be computed by the composite \eqref{eq:composite_intro_o2} in $\cat{L}$, which is a $\nZ_2$-graded complex of vector spaces $Y^{\vee} \l X$ equipped with the action of a Clifford algebra. If $Y(t), X(u)$ are allowed to depend on parameters, the cut becomes a vector bundle
\be\label{eq:orynx}
\mathcal{Z}'(O_2) := Y^{\vee}(t) \l X(u)
\ee
over $\mathscr{M}_Y \times \mathscr{M}_X$ equipped with a differential and compatible Clifford action, such that
\be
Y^{\vee}(t) \l X(u) \Big\vert_{t=P,u=Q} \cong Y^{\vee}(P) \l X(Q)
\ee
for every point $(P,Q)$.

Since $\cat{L}$ and $\LG$ are equivalent, the two ways \eqref{eq:wuhe20},\eqref{eq:orynx} of reading the composite \eqref{eq:composite_intro_o2} must contain the same information. However, the fact that $\eqref{eq:orynx}$ is a finite rank vector bundle over $\mathscr{M}_Y \times \mathscr{M}_X$ while \eqref{eq:wuhe20} is a complex of (non-finitely generated) quasi-coherent sheaves explains why $\cat{L}$ is better suited to this kind of problem.

%We leave the details to the aforementioned sequel \cite{??}, but here is a simple example which already shows the nontriviality of viewing correlators as functionals of the moduli of the labels on defects. Consider the following bordism which is a morphism from the empty manifold $\emptyset$ to $O$, where $X$ denotes a matrix factorisation of $V(y) - W(x)$. The value of this diagram is an element
%\[
%\mathcal{Z}(M) \in \mathcal{Z}(O) \cong Jac(V)
%\]
%By virtue of the above discussion, we may think of $\mathcal{Z}(M)$ as a \emph{section} of the trivial vector bundle with fiber $Jac(V)$ over the scheme $\mathscr{M}(W,V)$ which parametrises $1$-morphisms from $W$ to $V$ in $\LG$. If this section is not identically zero, then $V$ is called a \emph{generalised orbifold} of $W$ \cite{??} and it is of great interest to know when this is the case. From what we have said, it is clear that this is essentially a question of the geometry of $\mathscr{M}(W,V)$.

TODO: discuss $A_\infty$ algebras. If we want to understand how minimal models vary with the coefficients, and where the jumps occur, we need an intermediate finite object over blah since minimal models only work well over a field.

\subsubsection{Semantics of logic}

In forthcoming work we will use the results of this paper to define a novel semantics of linear logic, in which formulas are interpreted by tuples $(\md{X}, \nC[x], W)$ where $\md{X}$ is a scheme and $W \in \nC[x]$ is a potential. Proofs in the logic are interpreted by certain correspondences between these pairs. To explain briefly why we are forced to use $\cat{L}$ rather than $\LG$, we will use the denotation of the Church numeral $\underline{2}$ (see \cite{murfet_ll}).

Let $x$ be a variable of the logic and suppose it has denotation $\llbracket x \rrbracket = (\Spec(\nC), \nC[x], W)$ in the semantics. Let $\mathscr{M} = \mathscr{M}(W,W)$ be the (indexed) affine scheme parametrising $1$-morphisms $X: W \lto W$ in $\cat{L}$, as discussed above. Then the semantics works as follows:
\begin{gather*}
\llbracket x \multimap x \rrbracket = \big(\Spec(\nC), \nC[x,x'], W(x) - W(x') \big)\,,\\
\llbracket {!}( x \multimap x ) \rrbracket = \big( \mathscr{M}, \nC, 0 \big)\,.
\end{gather*}
The proof $\underline{2}$ of the sequent ${!}(x \multimap x) \vdash x \multimap x$ has for its denotation a correspondence between $\llbracket {!}( x \multimap x ) \rrbracket$ and $\llbracket x \multimap x \rrbracket$, which concretely means that $\llbracket \underline{2} \rrbracket$ is a bundle over the scheme $\mathscr{M}$ whose fiber over a loop $X$, viewed as a point $[X] \in \mathscr{M}$, is the square $X \l X$. More precisely, $\llbracket \underline{2} \rrbracket$ is the bundle over $\mathscr{M} \times \Spec(\nC[x,x'])$ whose fiber over $([X], a)$ is
\be\label{eq:bundledifferentialcliffordproof}
\llbracket \underline{2} \rrbracket_{[X],a} = \left( X \l X \right) \otimes_{\nC[x,x']} \nC(a)\,.
\ee
For these definitions to work, the coefficients of the monomials in the matrix factorisation $X \l X$ need to be given explicitly as polynomial functions of the coefficients in $X$. This means we have to use $\cat{L}$ rather than $\LG$.

@@

The cut operation in $\cat{L}$ produces objects which are representations of Clifford algebras. We briefly recall the basic theory of these algebras and their modules; for a full discussion see \cite{friedrich}. For $n \ge 0$ the Clifford algebra $C_n$ is the associative $\mathbb{Z}_2$-graded $k$-algebra generated by odd elements $\gamma_1,\ldots,\gamma_n$ and $\gamma_1^\dagger, \ldots, \gamma_n^\dagger$ subject to the following relations
\be\label{eq:clifford_relations}
[\gamma_i, \gamma_j] = 0, \qquad [\gamma_i^\dagger, \gamma_j^\dagger] = 0, \qquad [\gamma_i, \gamma_j^\dagger] = \delta_{ij}\,.
\ee
for all $1 \le i, j \le n$. In this paper all our commutators are graded, $[a,b] = ab - (-1)^{|a||b|} ba$. We set $C_0 = k$. These Clifford algebras have essentially only one nontrivial representation, sometimes called the \emph{spinor representation} \cite[p.14]{friedrich}, defined as follows:

@@

\begin{lemma}\label{defn:spinorrep} The map $C_n \lto \End_k(S_n)$ defined by
\[
\gamma_i \mapsto \theta_i \wedge (-), \qquad \gamma_i^\dagger \mapsto \theta_i^*\, \lrcorner\, (-)
\]
is an isomorphism of $\mathbb{Z}_2$-graded $k$-algebras.
\end{lemma}

@@

\begin{definition}\label{defn:idempotent_e} For $n \ge 0$ let $e_n \in C_n$ denote the element
\[
e_n = \gamma_1^\dagger \cdots \gamma_n^\dagger \gamma_n \cdots \gamma_1\,.
\]
This is the idempotent corresponding to the summand $k \cdot 1$ of $S_n$.
\end{definition}

@@

\section{The superbicategory $\cat{L}$}\label{section:lg_cut_system}

>>>>

% Operators
\def\eval{\operatorname{ev}}
\def\res{\operatorname{Res}}
\def\Coker{\operatorname{Coker}}
\def\Ker{\operatorname{Ker}}
\def\im{\operatorname{Im}}
\def\can{\operatorname{can}}
\def\K{\mathbf{K}}
\def\D{\mathbf{D}}
\def\N{\mathbf{N}}
\def\LG{\mathcal{LG}}
\def\Ab{\operatorname{Ab}}
\def\Hom{\operatorname{Hom}}
\def\modd{\operatorname{mod}}
\def\Modd{\operatorname{Mod}}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\nN{\mathds{N}}
\def\nZ{\mathds{Z}}
\def\nQ{\mathds{Q}}
\def\nR{\mathds{R}}
\def\nC{\mathds{C}}
\def\L{\mathcal{C}}
\def\ferm{\gamma}
\def\fermc{\gamma^\dagger}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\tot}{Tot}
\DeclareMathOperator{\ch}{ch}
\DeclareMathOperator{\str}{str}
\DeclareMathOperator{\hmf}{hmf}
\DeclareMathOperator{\HMF}{HMF}
\DeclareMathOperator{\hf}{HF}
\DeclareMathOperator{\At}{At}
\DeclareMathOperator{\Cat}{Cat}
\DeclareMathOperator{\Spec}{Spec}

@@

\begin{abstract}
The bicategory $\LG$ of Landau-Ginzburg models has polynomials as objects and matrix factorisations as $1$-morphisms. The composition of $1$-morphisms naturally produces infinite rank matrix factorisations, which is a nuisance. In this paper we define a bicategory $\L$ which is equivalent to $\LG$ but for which composition of $1$-morphisms produces finite rank matrix factorisations equipped with the action of a Clifford algebra.
\end{abstract}

@@

\section{Introduction}
% In the functorial approach to two-dimensional topological field theory with defects \cite{??,??} the data of the field theory is encoded as a bicategory with additional structure, including adjoints for all its $1$-morphisms. In the example of topological sigma models the associated bicategory has smooth projective varieties as objects and Fourier-Mukai kernels as $1$-morphisms \cite{??} while for topological Landau-Ginzburg models the associated bicategory $\LG$ has

The bicategory $\LG$ of Landau-Ginzburg models has a rich structure, as can be seen by its important role in topological field theory \cite{McNameethesis,lgdual,genorb} homological knot invariants \cite{kr0401268} and $2$-representations of quantum groups \cite{??}, but the definition is still relatively elementary. This makes $\LG$ an attractive algebraic object of study, but for one major wrinkle: the composition operation for $1$-morphisms is infinitary.

The objects of $\LG$ are potentials $W(x)$, while $1$-morphisms $W(x) \lto V(y)$ are matrix factorisations of the difference $V(y) - W(x)$ over $\nC[x,y]$. Such a matrix factorisation is a $\nZ_2$-graded free $\nC[x,y]$-module $X = X^0 \oplus X^1$ together with an odd operator $d_X$ satisfying $d_X^2 = (V(y) - W(x)) \cdot 1_X$. For a third potential $U(z)$, the composition of $1$-morphisms is a functor
\begin{gather*}
\LG(V,U) \times \LG(W,V) \lto \LG(W,U)\,,\\
(Y , X) \mapsto Y \circ X
\end{gather*}
% In this paper and its sequel we study an enrichment of $\LG$ which we refer to as a \emph{cut system}.
defined for matrix factorisations $Y$ of $V - U$ and $X$ of $W - V$ by
\be
Y \circ X = \big( Y \otimes_{\nC[y]} X, d_Y \otimes 1 + 1 \otimes d_X \big)\,.
\ee
This is the analogue of convolution of Fourier-Mukai kernels in the matrix factorisation setting and suffers from the same defect, namely, the resulting matrix factorisation $Y \circ X$ is a free module of \emph{infinite rank} over $\nC[x,z]$ which is only homotopy equivalent (that is, isomorphic in the homotopy category) to a finite rank matrix factorisation. While this finite model can be proven abstractly to exist, there is no obvious way to describe it as a function of the input data $Y,X$ viewed as matrices of polynomials. This need to deal throughout with infinite rank matrix factorisations introduces a certain minimal level of complexity into the theory, which for many applications is undesirable.

Together with Dyckerhoff we initiated in \cite{dm1102.2957} the study of these finite models in terms of homological perturbation and Atiyah classes, and in the present paper this program reaches its logical conclusion. We introduce a bicategory $\L$ in which the objects are again potentials and where the $1$-morphisms are finite rank matrix factorisations equipped with the structure of a representation of a Clifford algebra. We will prove that this bicategory is equivalent to $\LG$, but is much simpler - in particular, it has a composition operation which is described by a finite number of polynomial functions of the input data.

More precisely: the composition of $1$-morphisms in $\L$ is a functor
\begin{gather}
\L(V,U) \times \L(W,V) \lto \L(W,U)\,,\\
(Y,X) \mapsto Y \l X
\end{gather}
which we call the \emph{cut operation}. The definition of the cut is very simple. With $J_V = \nC[y]/(\partial V)$ the Jacobi algebra of $\nC$ (which is a finite-dimensional $\nC$-vector space by the assumption that $V(y)$ is a potential) the cut of $Y$ and $X$ is a finite rank matrix factorisation
\be
Y \l X = Y \otimes_{\nC[y]} J_V \otimes_{\nC[y]} X = ( Y \otimes_{\nC[y]} X ) / (\partial V) (Y \otimes_{\nC[y]} X )
\ee
over $\nC[x,z]$

equipped with odd operators $\ferm_i, \fermc_i$ satisfying Clifford relations
\begin{gather*}
\ferm_i e = \ferm_i = e \ferm_i\,, \\
\fermc_i e = \fermc_i = e \fermc_i\,.
\end{gather*}
up to homotopy (see Theorem \ref{theorem:cut_is_rep}). The differential on $Y \l X$ is induced from the differential on $Y \otimes X$ and can be written explicitly as a matrix over $\nC[x,z]$ in terms of the polynomial entries of $d_Y, d_X$. The Clifford action and the matrices which define the Clifford actions on $Y,X$. In particular, infinite rank matrix factorisations are never involved. In the remainder of the introduction we sketch how the cut operation arises from homological perturbation, why Clifford algebras are involved, where the equivalence $\L \cong \LG$ comes from, and our motivations for pursuing this ``finite model'' $\L$ of $\LG$. %In this sense the bicategory $\L$, which we prove is equivalent to $\LG$, provides a ``finite model'' of $\LG$. This has various applications, which we will describe later in the introduction.

The main theorem (Theorem \ref{theorem:htpy_equivalence_main}) says
\begin{equation}\label{eq:final_finite_model}
\xymatrix@C+4pc{
Y \l X \ar@<-1ex>[r]_-{\Phi^{-1}} & S_m \otimes_k ( Y \otimes X )\ar@<-1ex>[l]_-{\Phi}
}
\end{equation}
The paper is structured as follows: in Section \ref{section:background} we recall the definition of superbicategories and Clifford algebras, and introduce the Clifford thickening of a supercategory. In Section \ref{section:lg_cut_system} we define $\L$ and in Section \ref{section:equivllg} we prove it is equivalent to $\LG$ as a superbicategory without units. We do not consider unit $1$-morphisms here because $\L$ does not make working with units any simpler.

\subsection{$A_\infty$-algebras and minimal models}

The relationship between $\LG$ and $\L$ is similar to the relationship between a DG-algebra and its $A_\infty$-minimal model. Given a DG algebra $A$ over a field $k$, which is infinite-dimensional but has finite-dimensional cohomology, the $A_\infty$-minimal model is an $A_\infty$-structure on the cohomology $H^*(A)$ which is $A_\infty$-quasi-isomorphic to $A$. The tradeoff being made is that $A$ is infinite-dimensional but with a simple structure (i.e. one multiplication map $m_2$) while $H^*(A)$ is finite-dimensional but with possibly an infinite family of higher multiplications. In $\LG$ the composition operation is not described by a finite amount of data, and the price that one pays by going to $\L$ where this is the case, is that the composition has additional structure (a Clifford action). 
There is a more precise link to $A_\infty$-minimal models. A special case of the equivalence between $\L$ and $\LG$ gives, for a matrix factorisation $X$ of the potential $W(x_1,\ldots,x_m)$ viewed, a strong deformation retract
\be
\Phi: S_m \otimes_k \Hom_R(X,X) \lto X^{\vee} \l X\,.
\ee
where $S_m = \bigwedge( k \theta_1 \oplus \cdots \oplus k \theta_m )$ is an exterior algbera, $\Hom_R(X,X)$ is the Hom complex. This deformation retract can be used as the input data for the usual construction of an $A_\infty$-minimal on the left hand side, viewed as a DG-algebra.

\subsection{A geometric enrichment of $\LG$}

A finite rank matrix factorisation of $V(y) - W(x)$ is a matrix of polynomials
\[
D \in M_{2r}( \nC[x,y] ) \qquad \text{ satisfying } \qquad D^2 = (V(y) - W(x)) \cdot I_{2r}
\]
where we require that, dividing $D$ into $r \times r$ blocks, the upper left and bottom right blocks are zero matrices. The set of finite rank matrix factorisations is denoted $\LG(W,V)_{fin}$. For technical reasons the category of $1$-morphisms $\LG(W,V)$ contains, in addition to the finite rank matrix factorisations, those infinite rank matrix factorisations which are homotopy equivalent to finite rank ones; this means in particular that the inclusion
\[
\LG(W,V)_{fin} \lto \LG(W,V)
\]
is an equivalence of categories. If we fix the size $r$ of the blocks and bound the degrees of the monomials appearing in $D$ by some integer $k$, we can parametrise the possible matrix factorisations by the closed points of an affine scheme $\mathscr{M}_{r,k}(W,V)$ whose coordinates are the coefficients of the various monomials in each row and column of $D$. Hence as sets we have
\begin{equation}\label{eq:param_frmf}
\LG(W,V)_{fin} = \bigsqcup_{r \ge 1} \bigcup_{k \ge 0} \mathscr{M}_{r,k}(W,V)_{\nC}\,.
\end{equation}

does indeed lift to a morphism of the parametrising schemes - fulfilling the original goal. The schemes parametrising $1$-morphisms in $\L$ together with the morphisms of schemes lifting composition, and vector bundles encoding the $2$-morphisms, is what we refer to as a \emph{cut system} in the sequel \cite{cutsystems2}. However the present paper is concerned only with defining the superbicategory $\L$ and proving that it is equivalent to $\LG$. TODO: A cut system is a bicategory enriched in a certain sense in the category of (indexed families of) affine schemes.

%For various applications, two of which we sketch below, we would like to know that the basic operations of the bicategory $\LG$ respect this additional scheme structure; that is, that these operations are continuous in the coordinates which parametrise the $1$-morphisms. The most important and nontrivial such structure is the composition functor
%\begin{gather*}
%\LG(V,U)_{fin} \times \LG(W, V)_{fin} \lto \LG(W,U)\,,\\
%(Y,X) \mapsto Y \circ X\,.
%%\end{gather*}
%(finite model) We would like therefore to produce a morphism of schemes (omitting indices)
%\[
%c: \mathscr{M}(V,U) \times \mathscr{M}(W,V) \lto \mathscr{M}(W,U)
%\]
%with the property that $c([Y], [X]) = [Y \circ X]$, where $[-]$ denotes the point of the parameter space corresponding to a given matrix factorisation.

% The problem is that this doesn't make sense, because the matrix of polynomials defining the composite $Y \circ X$ contains monomials $x^i y^j z^k$ with $y$'s in them and we have no way to store the coefficients of such monomials in the space $\mathscr{M}(W,U)$. However one can prove that $Y \circ X$ is a summand in the homotopy category of a finite rank matrix factorisation (involving only $x$'s and $z$'s) and even describe this object and the corresponding idempotent explicitly \cite{dm1102.2957}. This is a finite amount of data which can be used to encode $Y \circ X$, so we just need to replace $\mathscr{M}(W,U)$ by some scheme which knows how to store such data.

\subsection{Semantics of linear logic}

In forthcoming work we will use the results of this paper to define a novel semantics of linear logic, in which formulas are interpreted by tuples $(\md{X}, \nC[x], W)$ where $\md{X}$ is a scheme and $W \in \nC[x]$ is a potential. Proofs in the logic are interpreted by certain correspondences between these pairs. To explain briefly why we are forced to use $\L$ rather than $\LG$, we will use the denotation of the Church numeral $\underline{2}$ (see \cite{murfet_ll}).

Let $x$ be a variable of the logic and suppose it has denotation $\llbracket x \rrbracket = (\Spec(\nC), \nC[x], W)$ in the semantics. Let $\mathscr{M} = \mathscr{M}(W,W)$ be the (indexed) affine scheme parametrising $1$-morphisms $X: W \lto W$ in $\L$, as discussed above. Then the semantics works as follows:
\begin{gather*}
\llbracket x \multimap x \rrbracket = \big(\Spec(\nC), \nC[x,x'], W(x) - W(x') \big)\,,\\
\llbracket {!}( x \multimap x ) \rrbracket = \big( \mathscr{M}, \nC, 0 \big)\,.
\end{gather*}
The proof $\underline{2}$ of the sequent ${!}(x \multimap x) \vdash x \multimap x$ has for its denotation a correspondence between $\llbracket {!}( x \multimap x ) \rrbracket$ and $\llbracket x \multimap x \rrbracket$, which concretely means that $\llbracket \underline{2} \rrbracket$ is a bundle over the scheme $\mathscr{M}$ whose fiber over a loop $X$, viewed as a point $[X] \in \mathscr{M}$, is the square $X \l X$. More precisely, $\llbracket \underline{2} \rrbracket$ is the bundle over $\mathscr{M} \times \Spec(\nC[x,x'])$ whose fiber over $([X], a)$ is
\be\label{eq:bundledifferentialcliffordproof}
\llbracket \underline{2} \rrbracket_{[X],a} = \left( X \l X \right) \otimes_{\nC[x,x']} \nC(a)\,.
\ee
For these definitions to work, the coefficients of the monomials in the matrix factorisation $X \l X$ need to be given explicitly as polynomial functions of the coefficients in $X$. This means we have to use $\L$ rather than $\LG$.

@@

The cut operation in $\L$ produces objects which are representations of Clifford algebras. We briefly recall the basic theory of these algebras and their modules; for a full discussion see \cite{friedrich}. For $n \ge 0$ the Clifford algebra $C_n$ is the associative $\mathbb{Z}_2$-graded $k$-algebra generated by odd elements $\ferm_1,\ldots,\ferm_n$ and $\fermc_1, \ldots, \fermc_n$ subject to the following relations
\be\label{eq:clifford_relations}
[\ferm_i, \ferm_j] = 0, \qquad [\fermc_i, \fermc_j] = 0, \qquad [\ferm_i, \fermc_j] = \delta_{ij}\,.
\ee
for all $1 \le i, j \le n$. In this paper all our commutators are graded, $[a,b] = ab - (-1)^{|a||b|} ba$. We set $C_0 = k$. These Clifford algebras have essentially only one nontrivial representation, sometimes called the \emph{spinor representation} \cite[p.14]{friedrich}, defined as follows:

@@

\begin{lemma}\label{defn:spinorrep} The map $C_n \lto \End_k(S_n)$ defined by
\[
\ferm_i \mapsto \theta_i \wedge (-), \qquad \fermc_i \mapsto \theta_i^*\, \lrcorner\, (-)
\]
is an isomorphism of $\mathbb{Z}_2$-graded $k$-algebras.
\end{lemma}

@@

\begin{definition}\label{defn:idempotent_e} For $n \ge 0$ let $e_n \in C_n$ denote the element
\[
e_n = \fermc_1 \cdots \fermc_n \ferm_n \cdots \ferm_1\,.
\]
This is the idempotent corresponding to the summand $k \cdot 1$ of $S_n$.
\end{definition}

@@

\section{The superbicategory $\L$}\label{section:lg_cut_system}
